---
title: "PCRNormalizator v1.0"
author: "RDA and MB"
date: "5/11/2021"
output: html_notebook
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Data loading:
The input data are:

- the matrix with Ct values for each miRNA (row) and each sample (column) 
- a sample metadata file, i.e. a tab delimited txt file with first column of sample names (matching Ct matrix's sample names) and second column with group specifications
See example files into data folder.

```{r raw data loading, echo=FALSE}
message('Please choose Ct matrix (.txt) to load:')
rawMirna <- read.table(file.choose(),
                       header = TRUE,
                       row.names = 1,
                       stringsAsFactors = FALSE)

message('Please add sample metadata to load:')
samples <- read.table(file.choose(),
                      header = TRUE,
                      stringsAsFactors = TRUE)
  
```

### Step1: parsing input data for NormqPCR and load it

```{r parsing input for NormqPCR packg}
library(reshape2)

write.table(melt(cbind(rawMirna, Detector =row.names(rawMirna)),
            id.vars       = "Detector",
            variable.name = "Sample",
            value.name    = "Cq")[c(2,1,3)], 
            file = "raw.4NormqPCR.txt",
            sep = "\t")


```



```{r load qPCR data, message=FALSE, warning=FALSE, include=FALSE}
#BiocManager::install("NormqPCR")
library(NormqPCR)

qPCR.data <- read.qPCR("raw.4NormqPCR.txt", 
                       phenoData = as(rawMirna, "AnnotatedDataFrame"))
file.remove("raw.4NormqPCR.txt")
```


### Step2: run geNorm

```{r run geNorm, echo=TRUE, message=TRUE, warning=FALSE}
geNorm.NormqPCR.Res <- selectHKs(qPCR.data, method = "geNorm",
                                   minNrHKs = 2, log =FALSE, 
# per l'opzione log, vedi https://www.researchgate.net/post/geNorm_in_NormqPCR_package_of_R
                                   Symbols = featureNames(qPCR.data),na.rm=TRUE)


```

### Step3: run Normfinder

```{r run Normfinder, include=FALSE}
res.normfinder <- stabMeasureRho(replaceNAs(qPCR.data, newNA = 40), 
                                 group = samples$group, 
                                 log = FALSE,returnAll=TRUE)

```

### Step4: calculate cv score
```{r calculate cv score}
cv.sample <- apply(rawMirna,2,function(x) sd(x, na.rm=TRUE)/mean(x, na.rm=TRUE))

cv.score <- apply(rawMirna,1,function(x) sd(x, na.rm=TRUE)/
                    mean(x, na.rm=TRUE))/
                    sum(cv.sample)
```



### Step5: combining all scores and calculating final Summarized Stability Score (SSS)
It produces stabilityScore.txt file


```{r combine stability scores}
df.rank <- Reduce(function(d1, d2) merge(d1, d2, by = "mirna",all=TRUE),
       list(data.frame("mirna"=gsub("\\.","-",geNorm.NormqPCR.Res$ranking),
                    "geNorm.meanM"=c(sort(geNorm.NormqPCR.Res$meanM),geNorm.NormqPCR.Res$meanM[1])), 
            data.frame("mirna"=gsub("\\.","-",names(res.normfinder$rho)),
                       "normFinder.rho"=res.normfinder$rho), 
            data.frame("mirna"=row.names(rawMirna),
                       "cv.score"=cv.score)))        

            
            
df.rank$SSS <- sqrt(df.rank$geNorm.meanM^2+df.rank$normFinder.rho^2+df.rank$cv.score^2)
df.rank[-1] <- round(df.rank[-1],digits = 4)

df.rank.F <- df.rank[order(df.rank$SSS),]


write.table(df.rank.F,file="stabilityScore.txt",
            sep = "\t",quote = FALSE,row.names = FALSE)

```


### Step6: plot scores for 10 most stable mirnas according to the SSS
Plots are saved in stabilityScore.pdf
```{r plots, echo=FALSE}
library(ggpubr)
library(gridExtra)


plots <- list()


plots[[1]] <- ggline(df.rank.F[1:10,], x = "mirna", y = "geNorm.meanM",color = "blue",title = "geNorm",
                     ylab = "mean M") + border("black") + grids(linetype = "dashed") + 
  rotate_x_text(90) + rremove("xlab") + font("ylab", size = 14)

plots[[2]] <- ggline(df.rank.F[1:10,], x = "mirna", y = "cv.score",color = "orange",title = "CV score",
                     ylab = "CV score") + border("black") + grids(linetype = "dashed") + 
  rotate_x_text(90) + rremove("xlab") + font("ylab", size = 14)

plots[[3]] <- ggline(df.rank.F[1:10,], x = "mirna", y = "normFinder.rho",color = "darkgreen",title = "NormFinder",
                     ylab = "rho") + border("black") + grids(linetype = "dashed") + 
  rotate_x_text(90) + rremove("xlab") + font("ylab", size = 14)

plots[[4]] <- ggline(df.rank.F[1:10,], x = "mirna", y = "SSS",color = "red",title = "Summarized Stability Score",
                     ylab = "SSS") + border("black") + grids(linetype = "dashed") + 
  rotate_x_text(90) + rremove("xlab") + font("ylab", size = 14)

pdf("stabilityScores.pdf",6,16)
grid.arrange(grobs = plots[1:4], ncol = 1)
dev.off()


```